A full approximation scheme multilevel method \\ for nonlinear variational inequalities

### 2023-07-15 paper/fascd.tex
draft a discussion and outlook section

Specifically, our \emph{full approximation scheme constraint decomposition} (FAScd) method extends the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} in the following five directions:
\renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
\begin{enumerate}
\item The FAS coarse corrections are adapted for general VI problems. The needed computations require neither an objective function nor linearity of the residual operator.
\item Box constraints, namely upper and lower obstacles, are permitted.
\item We implement V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}).
\item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
\item By using a Newton-type active set method as a smoother, with a fixed number of (Newton and Krylov) iterations per step, so that the work of the smoother is proportional to the number of unknowns on the given level, we demonstrate that F-cycle FAScd is an optimal solver for the examples tested (Section \ref{sec:results}).
\end{enumerate}

@@

FIXME For the classical obstacle problem, based on the linear, symmetric Laplacian operator, certain multilevel techniques, such as the truncated monotone multigrid method \cite{GraeserKornhuber2009,Kornhuber1994}, are known to improve performance relative to the original multilevel CD method \cite{Tai2003}.  These improved methods modify the nodal basis functions.  By contrast FAScd makes essentially no assumptions about the FE discretization.  Acceleration via basis-level manipulations is a topic for future research.

FIXME So the difference between FAScd and \cite{GraeserKornhuber2009} are: (1) FAS coarse corrections, (2) no attempt to do monotone/truncated multigrid which would be a mess (I think) across the FE infrastructure, (3) clarity on what sets you can use in $V(\nu_1,\nu_2)$ cycles, with bigger up-smoother sets in particular, and most importantly (4) implementation within the mature Firedrake/PETSc library framework where one can easily hook up an active-set Newton step as a smoother.

FIXME  It is hard to say which aspect of FAScd is more important to beating monotone multigrid: stronger smoother or improved up-smooth CD.  Probably the former, but for now we don't have any illuminating analysis, and a head-to-head race is comparing apples and oranges because we are using such a strong smoother.  Various differences of implementation don't produce any really good multigrid convergence if your smoother is NGS on each level.  I think the convergence proofs in \cite{GraeserKornhuber2009} all depend on this?

FIXME the fact that FAScd reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}) suggests it is a completely different strategy from PFAS in \cite{BrandtCryer1983}

FIXME main difference relative to monotone multigrid \cite{Kornhuber1994} and PFAS \cite{BrandtCryer1983} is that method is not tied to, or presented in terms of, an PNGS type smoother

FIXME at a high level one might say FAScd merely generates coarse VI problems which, once solved, make helpful admissible additions to the eventual V-cycle update; the method is smoother-agnostic; smoother could be Newton semi-smooth \cite{BensonMunson2006}, interior point method, etc.

FIXME VINewton smoother could use GMG or AMG on linear step equations, but this was not seen to be necessary in applications so far and was not even tested

FIXME the observation that up-smoothing is more efficient, with consequent asymmetry between $V(1,0)$ and $V(0,1)$ performance, seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}

FIXME desired extension to nonlocal variational inequalities not tried here

>>>>

Specifically, our \emph{full approximation scheme constraint decomposition} (FAScd) method extends the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} in the following five directions:
\renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
\begin{enumerate}
\item The FAS coarse corrections are adapted for general VI problems. The needed computations require neither an objective function nor linearity of the residual operator.
\item Box constraints, namely upper and lower obstacles, are permitted.
\item We implement V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}).
\item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
\item By using a Newton-type active set method as a smoother, with a fixed number of (Newton and Krylov) iterations per step, so that the work of the smoother is proportional to the number of unknowns on the given level, we demonstrate that F-cycle FAScd is an optimal solver for the differential VI examples in Section \ref{sec:results}.
\end{enumerate}

@@

As a multilevel solver for VI problems, one might describe the new FAScd method as merely a strategy for generating coarser-level problems which, once solved, make helpful admissible additions to a final V-cycle update.  The method is smoother-agnostic, and our implementation within the extensible Firedrake \cite{Rathgeberetal2016} and PETSc \cite{Balayetal2023} library framework should allow easy experimentation with other smoothers (projected Gauss-Seidel, semi-smooth Newton, interior-point, etc.), beyond the active-set Newton method \cite{BensonMunson2006} chosen for demonstration here.  For box-constrained smoothers of Newton-Krylov type one could also add linear geometric or algebraic multigrid preconditioning \cite{Trottenbergetal2001}, instead of our simple choice of incomplete factorization and (in parallel) domain decomposition preconditioners.  Thus there is a large smoother and coarse solver space to explore.  However, performance of the particular method shown in Section \ref{sec:results} is already quite good.

The observation that up-smoothing is more efficient in a multilevel CD context, which arises because of constraint set distinctions, seems to be new, but compare comments on $V(1,1)$ cycles in \cite{GraeserKornhuber2009,Tai2003}.  The increased efficiency is shown in Example \ref{ex:results:plap} particularly.  In summary, the distinction is that construction of constraint sets via monotone injection generates up-smoothing sets which can telescope the down-smoothing sets (Lemma \ref{lem:upwardadmissibility}); we call this an \emph{incomplete CD}.  This structure can be exploited by a new CD iteration, algorithm \pr{icd-tele} in Section \ref{sec:cdmultilevel}.

In contrast with monotone multigrid \cite{Kornhuber1994}, FAScd does not modify the FE discretization, but the possibility of acceleration via basis-level manipulations is a topic for future research.  On the other hand,  the convergence proofs in \cite{GraeserKornhuber2009,Kornhuber1994}, which appear to depend on particular smoother and basis-level manipulation choices, have no analog here.

%While an an extension to nonlocal, non-differential VI problems is of interest, it was not tried here
